\chapter{Introduction}
\section{Overview}
Magic: The Gathering (MTG) is a Trading Card Game (TCG) designed by Wizards of the Coast (WOTC). Players take the role of a wizard, whose deck is a library of spells with which they battle one or more similarly equipped opponents. Pursuit of the hobby thus involves mastery of both gameplay and deck construction. While tournament-level participation relies on ``netdecking''~---~copying a decklist with a history of competitive success~---~the hobby is intended to have a significant creative component, with the unranked Commander format gaining popularity in recent years through its focus on unique and personal decks~\cite{CommanderFormatOverview}.

In addition to spell cards, decks also contain ``land'' cards, which generate ``mana'', a resource expended to cast spells. A deck's lands collectively form its ``manabase''. Mana comes in five colours: White, Blue, Black, Red and Green (abbreviated respectively to W, U, B, R and G, or WUBRG collectively). Spells require specific colours, and lands produce one or more colours. A deck including spells of many colours has access to more spells, but runs an increased risk of a ``colour screw'', in which the player does not draw lands of the colours they need. Whereas selecting a list of spells is a stimulating creative pursuit, choosing a deck's manabase is less so: within a set budget, and notwithstanding the minority of lands which come with effects outside mana provision, there are objectively lists of lands that will maximize a player's chances of being able to play their spells. 

LandFill is a webapp that automates this aspect of deck building. It models the selection of lands for a deck as a combinatorial optimization problem, where a given manabase yields a measurable performance score for the deck as a whole. This performance score is necessarily approximate: Churchill \textit{et al} have demonstrated that, as it is possible to construct within an MTG game a Turing Machine whose halting is the necessary condition for a player's victory, a deck's winning strategy is undecidable \cite{churchill2019magic}, meaning that it is beyond LandFill's capacity to determine what lands support the most decisive plays. Nevertheless, the heuristic that the winning MTG player is typically the one who spends the most mana over the course of the game \cite{KarstenCurve} provides an opening for approximating a perfromance score via Monte Carlo methods. Using a stripped-down MTG simulator, in which the simulated player aims only to spend as much mana as possible each turn, LandFill estimates over iterated games how effective a given manabase is. It then produces successively optimized manabases via a Hill Climbing Algorithm. Via a web interface, users are then able to input a deck's spell cards into this system and receive a completed decklist that had been broadly optimized within their set preferences. Initial user testing suggests that users are able to use LandFill to generate manabases of a comparable or greater quality to those they can assemble themselves in less time and with less effort. 

\section{Thesis Layout}
My writeup here documents the development and testing of LandFill. I will begin by outlining relevant MTG rules and design trends, and the difficulties these introduce for optimization, followed by an overview of how LandFill will approach these. I will then outline my initial consultations with MTG players, and the features they require from a manabase optimization app. Dividing LandFill into four components~---~a MTG card database, an MTG game simulator, an optimization algorithm implementer, and a user interface~---~I will then outline its structure, and justify the decisions made during development. I will then summarize a second round of user-testing, and outline the value added by LandFill. I will conclude with an evaluation of the strengths and limitations of the app, and areas for future development.